\chapter*{Preface}
% First sentence statement
The presented dissertation emerged from my work as a researcher with the Department of Communications Engineering at the University of Bremen.
% Thanks to "Doktorvater"
First and foremost I would like to thank Armin Dekorsy for offering me the opportunity to pursue my doctoral degree at the Department of Communications Engineering.
His continuous support and professional guidance enabled and encouraged me to successfully publish my results in numerous international conferences as well as to discuss and exchange ideas among peers to grow my scientific expertise as well as to grow personally.

%Thanks for examiners and further profs
Furthermore, I would like to extend my gratitude to Peter Rost for taking the time to assess this dissertation and for his valuable remarks.
The same goes for Steffen Paul and Karl-Ludwig Krieger for spending their time to read through this work and partaking in the colloquium as further examiners.
% Thanks to Jondral for his networking support
I would like to thank Friedrich K. Jondral for encouraging and supporting me to start this journey to pursue my doctorate.
% Thanks for the moniez!
I would further like to thank the German Federal Ministry of Education and Research (BMBF) for funding
a large portion of this work as part of the projects HiFlecs, TACNET 4.0, and IRLG.

% Thanks to Carsten
I am indebted to Carsten Bockelmann for his guidance, broad wisdom and skillful support during these years that made this thesis possible.
The countless discussions as well as support and time that enabled me to bring new ideas from inception to publication were immeasurably valuable.
% Thanks to coworkers
My coworkers deserve praise for making this time with the ANT as fruitful and enjoyable as anyone can only wish for.

% Thanks for friends, family, AND girlfriend
This multi year effort would not be possible without my friends and family who were always there for me.
I am deeply grateful for their support.
Finally, I would like to express my deepest gratitude to my girlfriend who supported me in every imaginable way to follow through with my work.
\\~\\
Bremen, Month 20XX\\
\\
Johannes Demel
